Credit risk distribution and capital savings

### 2019-08-19 Analysis/Draft/Draft.Rmd
First version of paper structure with to-do list

---
title: "Distribution of portfolio credit risk and capital savings"
author:
- affiliation: Tilburg University; l.avezum@uvt.nl
  name: Lucas Avezum
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
  html_document:
    df_print: paged
  word_document: default
bibliography: null
citecolor: null
anonymous: no
fontsize: 12pt
geometry: margin= 1in
keywords: Key1, Key2, Key3
linkcolor: blue
biblio-style: apsr
subtitle: ''
thanks: Very preliminary draft
#abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
#  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
#  a concave mapping from probabilities of default to capital requiremets. As a consequence,
#  banks can reduce requirements or increase risk-taking by widening their portfolio
#  credit risk distribution.'
toc: no
urlcolor: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = TRUE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

@@

library(tidyverse)           ## Data manipulation, pipe operator
library(splitstackshape)     ## Command to expand data
library(data.table)          ## Command to bind lists
library(ggridges)            ## Density ridges plot
library(stargazer)           ## Latex tables
library(lfe)                 ## High-dimensional fixed effects
library(gridExtra)           ## Plot side-by-side
library(DescTools)           ## Command to winsorize variables
library(robustHD)

p_w <- c(0.01, 0.99)         ## Set winsorizing limits
```

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
load("../../Data/Datasets/DifInDif.Rda")
```

@@

bank.data <- bank.data %>% ungroup() %>%
  mutate(RW.hat=RW.hat*100,
         RW.s=RW.s*100)
```

# List
* Use the share under IRB to weight capital savings measure (share vs intensity). Include share in regressions plus interaction with capital savings

* Explore alternative capital savings measures, such as, savings in terms of capital (mn),in terms of capital ratio, in terms of capital ratio savings (counterfactual vs actual)

* Plot marginal increase of savings in time.

* Decompose capital savings: share effect, size effect, etc...

* Remove step effect from regressions: average across all IRB banks (add IRB dummy) or IRB specific dummy (average out the step for each bank). Idea is to focus on difference across IRB banks. it might not make sense to include SA banks anymore.

* Plot capital savings measures to other risk variables

* Plot cross section for different capital saving measures (replicate first plot)

* check and better clean data:

# Capital savings measure

@@

\begin{equation}
\text{Capital savings}_{i}= \frac{f(\overline{PD})\sum_{i}EAD_iLGD_i-\sum_{i}EAD_iLGD_if(PD_i)}{\sum_{i}EAD_i},
\end{equation}

@@

\begin{equation}
\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
\end{equation}

```{r}
savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation 1. Each point in the plot is a bank-year observation."
```

```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
 ggplot(mapping = aes(x = reorder(name, RW.s, FUN = median, na.rm=TRUE),
                          y = RW.s,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital savings") +

  theme_bw()

@@

@@

# Decomposition

\begin{equation}
\begin{aligned}
\Delta RW & = q_{B}^{SA} \Delta RW^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \hat{RW}^{j} \\ 
& + (1-q_{B}^{SA})\sum_{j=1}^{P} \hat{RW}^{j} \Delta q^{j} -(1-q_{B}^{SA})\Delta RW^{S} + (RW^{SA}-RW^{IRB})\Delta q^{SA},
\end{aligned}
\end{equation}

\newpage
## Time decomposition

@@

```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}
time.lable <- c("t=0", "t=1", "t=2", "t=3", "t=4")
plot1 <-time.decomposition %>% 
  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain"),
         RW = RW - t0.RW,
         time = as.factor(time)) %>%
  ggplot(mapping = aes(x = time,
                       y    = value, 
                       fill = condition)) +
geom_bar(stat = 'identity')+
geom_line(aes(group=1, y = RW))  +
  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
  scale_x_discrete( name=NULL,breaks = c("0","1","2","3","4"),labels = c("t=0","t=1","t=2","t=3","t=4")) +
 theme(legend.position = c(0.3, 0.2), 
        legend.title = element_blank(), 
        legend.background=element_blank())

year.decomposition <- year.decomposition %>%
  mutate(RW_2 = RW - t0.RW,
         year = as.factor(year)) 
plot2 <- year.decomposition %>% 
  select(year,RW,  SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = year)) +
geom_bar(aes(y = value, 
          fill = condition),
          stat = 'identity')+
geom_line(data = year.decomposition, 
          aes(group=1,y = RW_2))  +
  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
theme(legend.position='none')
grid.arrange(plot1, plot2, ncol=2)

@@

## Cross-section decomposition

```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
plot1 <- cross.section.decomposition %>% 
  mutate(order.s = abs(dispersion.gain))%>%
  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = reorder(name, order.s),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  geom_line(aes(y = total,group=1))  +
  labs(x=NULL, y= NULL,subtitle = "Plot a") +
    coord_flip() +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
 theme(legend.position='none')

plot2 <- cross.section.decomposition %>% 
  mutate(order.s = abs(dispersion.gain))%>%
  select(name, Country, order.s, SA.gain, IRB.gain, mix.gain, dispersion.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(value = abs(value),
         condition = relevel(condition,"dispersion.gain")) %>%
  ggplot(mapping = aes(x = reorder(name, order.s),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity',position = "fill")+
  labs(x=NULL, y= NULL,subtitle = "Plot b") +
    coord_flip() +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
   theme(legend.title = element_blank(), 
        legend.background=element_blank(),
       legend.text=element_text(size=7),
       axis.text.y = element_blank())

# Counterfactual analysis

```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
 plot1 <- bank.data %>%  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>% 
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
theme(legend.position='none')

plot2 <- bank.data %>%  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="",subtitle = "Plot b") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  theme(legend.position = c(0.8, 0.85), 
        legend.title = element_blank(), 
        legend.background=element_blank())

plot3 <- bank.data %>%  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

plot4 <- bank.data %>% filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="",  subtitle = "Plot d") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  subset(select = c("RW.s","tier1ratio","tier.hat","totalcapitalratio","CAR.hat","ROE","Zscore")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("Capital savings","Tier 1 ratio","Counterfactual tier 1 ratio","Total ratio","Counterfactual total ratio","ROE","Z-score"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{11cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

@@

>>>>

---
title: "Credit risk distribution and capital savings"
author:
- name: Lucas Avezum
  affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan 2, 5037 AB Tilburg, NL, The Netherlands 
date: 
header-includes:
   - \usepackage{rotating, graphicx, dcolumn}
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: tex-ms.tex
  html_document:
    df_print: paged
  word_document: default
bibliography: ../../Auxiliar/mybibfile.bib
citecolor: black
anonymous: no
fontsize: 12pt
geometry: margin= 1in
keywords: Key1, Key2, Key3
linkcolor: blue
biblio-style: apsr
subtitle: ''
thanks: Very preliminary draft
#abstract: 'In this paper, I study the possibility, allowed by regulation, for banks
#  to increase their capital ratio while maintaning the average portfolio riskiness. Basel II introduced
#  a concave mapping from probabilities of default to capital requiremets. As a consequence,
#  banks can reduce requirements or increase risk-taking by widening their portfolio
#  credit risk distribution.'
toc: no
urlcolor: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = FALSE, 
                      warning   = FALSE, 
                      error     = FALSE, 
                      tidy      = TRUE, 
                      tidy.opts = list(width.cutoff=50), 
                      collapse  = TRUE,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE, 
                      comment   = "",
                      cache     = FALSE)
# These options are tuned for manuscript/presentation. 
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

@@

library(tidyverse)           ## Data manipulation, pipe operator
library(splitstackshape)     ## Command to expand data
library(data.table)          ## Command to bind lists
library(ggridges)            ## Density ridges plot
library(stargazer)           ## Latex tables
library(lfe)                 ## High-dimensional fixed effects
library(gridExtra)           ## Plot side-by-side
library(DescTools)           ## Command to winsorize variables

p_w <- c(0.01, 0.99)         ## Set winsorizing limits

```{r load data, include=TRUE}
load("../../Data/Datasets/Pillar3Data.Rda")
load("../../Data/Datasets/BankData.Rda")
load("../../Data/Datasets/CrossSectionDecomposition.Rda")
load("../../Data/Datasets/timedecomposition.Rda")
load("../../Data/Datasets/yeardecomposition.Rda")
```

@@

# To-Do List {-}

\begin{itemize}

\item Data collection

   \begin{itemize}

   \item[$-$] Collect information for 32 IRB banks that participate on EBA exercise out of 52. Currently, we have information for 10 banks (and 35 SA banks).

   \end{itemize}

\item Write data section:

   \begin{itemize}

   \item[$-$] Use decomposition to show how the intensity and extensity of PD distribution effect varies in time and across banks.

   \item[$-$] Use decomposition to calculate how much of the PD distribution effect variation explains total RW variation (both in time and across banks).

   \item[$-$] Explain capital savings measure

   \item[$-$] Summary statistics

   \end{itemize}

\item Analysis

   \begin{itemize}

   \item[$-$] EBA vs non-EBA:Implement matching strategy following \cite{gropp2018banks}.

   \item[$-$] Dif-on-Dif for EBA sample. Use capital shortfall as treatment intensity indicator.

   \item[$-$] OLS relating capital savings measure to, capital ratio, RWA, ROA, ROE, Z-score, leverage.

   \end{itemize}

\item Robustness
   \begin{itemize}

   \item[$-$] Previous analysis with alternative capital savings measures (Gini coefficient, RW)

   \end{itemize}

\item Extension

   \begin{itemize}

   \item[$-$] Use Barth measures to study interaction with regulation.

   \item[$-$] Use corporate governance \citep{anginer2016corporate} and subordinated debt \citep{gropp2018banks} to study different incentive to raise capital.

   \item[$-$] Use NPL ratios to check if PD dispersion is associated with better risk management practices [@cucinelli2018credit].

   \item[$-$] Check portfolio-level effects, for instance, corporate vs real estate.

   \end{itemize}
\end{itemize}
\newpage

# Introduction

# Background

@@

\begin{equation}
\label{eq1}
\text{RWA}= 12.5 \sum_{i}EAD_iLGD_i f(PD_i),
\end{equation}

\begin{equation}
\label{eq2}
\widehat{\text{RWA}}= 12.5 f(\overline{PD})\sum_{i}EAD_iLGD_i,
\end{equation}

@@

\begin{equation*}
\overline{PD}=\frac{\sum_iEAD_iLGD_iPD_i}{\sum_iEAD_iLGD_i}.
\end{equation*}

The capital savings measure is the difference between equations \ref{eq1} and \ref{eq2}:

\begin{equation}
\label{eq3}
\text{RWA savings}=\widehat{\text{RWA}}-\text{RWA},
\end{equation}

# Data

```{r}
savings_cap = "The figure plots the log of banks capital savings. This measure is calculated using equation \\ref{eq3}. Each point in the plot is a bank-year observation."
```

```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
 ggplot(mapping = aes(x = reorder(name, log.RWA.s, FUN = median, na.rm=TRUE),
                          y = log.RWA.s,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital savings (USD mn)") +
  theme_bw()

@@

## Capital Ratio Savings

\begin{equation}
\label{eq4}
\text{Counterfactual capital ratio}=\frac{\text{Capital}}{\widehat{\text{RWA}}},
\end{equation}

\begin{equation}
\label{eq5}
\text{Capital ratio savings}=\text{Counterfactual capital ratio}- \text{Observed capital ratio},
\end{equation}

```{r}
savings_cap = "The figure plots the capital ratio savings measure. This measure is calculated using equation \\ref{eq5}. Each point in the plot is a bank-year observation."
```

```{r, fig.height = 5, fig.width = 5,results="hide",fig.cap= savings_cap,fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), IRB == 1) %>%
 ggplot(mapping = aes(x = reorder(name, CAR.dif, FUN = median, na.rm=TRUE),
                          y = CAR.dif,
                      color= Country, 
                      fill = Country)) +
geom_point(alpha = 0.5,size = 2) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "Capital ratio savings (%)") +
  theme_bw()

```{r , fig.height = 6, fig.width = 7, results="hide",fig.cap= tier1_cap,out.width = "1\\textwidth"}
 plot1 <- bank.data %>%  ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>% 
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="Tier 1 capital ratio", subtitle = "Plot a") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"))+
  scale_color_discrete(labels = c("Actual", "Counterfactual")) +
theme(legend.position='none')

plot2 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, tier1ratio, tier.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, tier1ratio:tier.hat, factor_key=TRUE)  %>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="",y="",subtitle = "Plot b") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Tier 1 ratio")+
  theme(legend.position = c(0.8, 0.85), 
        legend.title = element_blank(), 
        legend.background=element_blank())

plot3 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="Total capital ratio", subtitle = "Plot c") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

plot4 <- bank.data %>% ungroup() %>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>% 
  mutate_if(is.numeric,Winsorize,probs = c(0.01, 0.99) ,na.rm = TRUE) %>%
  select(name,RW.s, totalcapitalratio, CAR.hat, IRB) %>%
  filter(!is.na(RW.s), IRB == 1) %>%
  group_by(name)%>%
 summarise_if(is.numeric,funs(mean),na.rm=TRUE)%>%
  gather(condition, value, totalcapitalratio:CAR.hat, factor_key=TRUE)%>%
 ggplot(mapping = aes(x = RW.s,
                      y = value,
                      color = condition,
                      fill = condition)) +
geom_point( alpha = 0.5,size = 1) +
  scale_y_continuous( limits=c(5, 20))+
  labs(x="Capital savings",y="",  subtitle = "Plot d") +
  geom_smooth(method = "lm",color="black") +
  theme_bw()+
  scale_color_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  scale_fill_discrete(labels = c("Actual", "Counterfactual"), name = "Total capital ratio")+
  theme(legend.position='none')

@@

## Decomposition

Rearranging \ref{eq3} and dividing by EAD we have:

\begin{equation}
\begin{aligned}
RW & =  \widehat{\text{RW}} +\text{RWA savings} \\
   & =  \widehat{\text{RW}} +\text{RW}^s 
\end{aligned}
\end{equation}

I use the method proposed by @cannata2012inside to decompose the components of risk-weights.

\begin{equation}
\begin{aligned}
\Delta \text{RW} & = q_{B}^{SA} \Delta \text{RW}^{SA}+(1-q_{B}^{SA})\sum_{j=1}^{P} q_{B}^{j}\Delta \widehat{\text{RW}}^{j} \\ 
& + (1-q_{B}^{SA})\sum_{j=1}^{P} \widehat{\text{RW}}^{j} \Delta q^{j}  + (\text{RW}^{SA}-\widehat{\text{RW}})\Delta q^{SA} -(1-q_{B}^{SA})\Delta \text{RW}^{S}+ \text{RW}^{S}\Delta q^{SA}\\
& = \text{Other effects} -\underbrace{(1-q_{B}^{SA})\Delta \text{RW}^{S}}_\text{Intensive margin}+ \underbrace{\text{RW}^{S}\Delta q^{SA}}_\text{Extensive margin}
\end{aligned}
\end{equation}

@@

```{r, fig.height = 4, fig.width = 7, results="hide",fig.cap= time.dec_cap,out.width = "1\\textwidth"}

plot1 <-time.decomposition %>% 
  select(time, RW, t0.RW,  SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain")),
         RW = RW - t0.RW,
         time = as.factor(time)) %>%
  ggplot(mapping = aes(x = time,
                       y    = value, 
                       fill = condition)) +
geom_bar(stat = 'identity')+
geom_line(aes(group=1, y = RW))  +
  labs( y = 'Change from t=0', subtitle = "Plot a" ) +
  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
  scale_x_discrete( name=NULL,breaks = c("0","2","4","6","8"),labels = c("t=0","t=2","t=4","t=6","t=8")) +
 theme(legend.position = c(0.3, 0.1), 
        legend.title = element_blank(), 
        legend.background=element_blank())

year.decomposition <- year.decomposition %>%
  mutate(RW_2 = RW - t0.RW,
         year = as.factor(year)) 
plot2 <- year.decomposition %>% 
  select(year,RW,  SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
  ggplot(mapping = aes(x = year)) +
geom_bar(aes(y = value, 
          fill = condition),
          stat = 'identity')+
geom_line(data = year.decomposition, 
          aes(group=1,y = RW_2))  +
  labs(y = 'Change from 2008', x=NULL, subtitle = "Plot b") +
  scale_fill_discrete(labels = c("PD distribution effect", "SA effect", "IRB effect","Mix effect","Roll-out effect"), name = "Component")+
  scale_x_discrete(breaks = c("2008","2010","2012","2014","2016","2018"))+
theme(legend.position='none')
grid.arrange(plot1, plot2, ncol=2)

```{r, fig.height = 5, fig.width = 7, results="hide",fig.cap= cross.dec_cap,out.width = "1\\textwidth"}
plot1 <- cross.section.decomposition %>% filter(q.SA<1)%>%
  mutate(order.s = abs(int.gain))%>%
  select(name, Country, total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
  ggplot(mapping = aes(x = reorder(name, total),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity')+
  geom_line(aes(y = total,group=1))  +
  labs(x=NULL, y= NULL,subtitle = "Plot a") +
    coord_flip() +
  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
 theme(legend.position='none')

plot2 <- cross.section.decomposition %>% filter(q.SA<1)%>%
  mutate(order.s = abs(int.gain))%>%
  select(name, Country,total= total.gain, order.s, SA.gain, IRB.gain, mix.gain, int.gain, ext.gain, rollout.gain) %>%
  gather(condition, value, contains(".gain"), factor_key=TRUE)%>%
  mutate(value = abs(value),
         condition = factor(condition,levels=c("int.gain","ext.gain"))) %>%
  ggplot(mapping = aes(x = reorder(name, total),
                       y = value, fill = condition)) +
geom_bar(stat = 'identity',position = "fill")+
  labs(x=NULL, y= NULL,subtitle = "Plot b") +
    coord_flip() +
  scale_fill_discrete(labels = c("Intensive effect", "Extensive effect", "Others"), name = "Component")+
   theme(legend.title = element_blank(), 
        legend.background=element_blank(),
       legend.text=element_text(size=7),
       axis.text.y = element_blank())

@@

```{r , fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
p_w <- c(0.01, 0.99)         ## Set winsorizing limits
bank.data %>% ungroup()%>%
  filter(!is.na(RW.s),
         !is.na(basel),
         tier1capitalmillcu < totalcapitalmillcu)%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  subset(select = c("log.RWA.s","tier1ratio","tier.hat","tier.dif","totalcapitalratio","CAR.hat","CAR.dif","ROE","ROA","Zscore","leverage")) %>%
  as.data.frame() %>% 
stargazer(header = FALSE, type="latex", single.row = TRUE, column.sep.width = "1pt", font.size = "small",
          omit.summary.stat = c("p25","p75"),
          covariate.labels = c("Log RWA savings","Tier 1 ratio","Counterfactual tier 1 ratio","Tier 1 ratio savings","Total ratio","Counterfactual total ratio","Capital ratio savings","ROE","ROA","Z-score","Leverage ratio"),
          title = "Summary statistics",
          table.layout = "n=d#-t-as-",
          notes = "\\parbox[t]{15cm}{The table shows summary statistics for the sample of banks collected. Capital savings, Portfolio mean PD and, Gini coefficient were calculated using the collected information from Pillar-3 reports. The remaining variables come from BankFocus.}", 
          notes.label = "",
          notes.align = "c", 
          notes.append = FALSE)
```

@@

# Results

pre.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2009","2010"),
         !is.na(IRB)) %>%
  select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)

post.treat.out <- bank.data %>%
  mutate(IRB = ifelse(IRB == 1, ifelse(EAD.IRB==0,NA,IRB),IRB))%>%
  filter(year %in% c("2012","2013"),
         !is.na(IRB)) %>%
  select(bvdid, log.RWA.total, log.RWA.other, log.RWA.credit, log.RWA.IRB, log.RWA.SA, log.RWA.hat, log.RWA.s,
         log.K, log.tier, totalcapitalratio, tier1ratio, CAR.dif, tier.dif) %>%
  group_by(bvdid) %>%
  summarize_if(is.numeric, funs(mean), na.rm = TRUE)
